Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-20 paper.tex
proved check consistency before compose middles;

@@

	\comprule{
		\hyperCoercionI
	}{
		\hyperCoercionC{\epsilon}{m}{t}
	}{
		\hyperCoercionC{?^{l}}{m}{t}
	}{\sidecond{\ell = l}}

@@

	\comprule{
		\hyperCoercionC{h}{m_1}{t_1}
	}{
		\hyperCoercionC{\epsilon}{m_2}{t_2}
	}{
		\hyperCoercionC{h}{m'}{t'}
	}{
		\sidecond{
			m_1 \fatsemi^{\ell} (m_2, t_2) = (m', t')
			 \; \text{and} \;
			 \forall l. t \neq \bot^{l}
		}
	}

	\comprule{
		\hyperCoercionC{h}{m_1}{t_1}
	}{
		\hyperCoercionC{?^{l'}}{m_2}{t_2}
	}{
		\hyperCoercionC{h}{m'}{t'}
	}{
		\sidecond{
			m_1 \fatsemi^{l'} (m_2, t_2) = (m', t')
			\; \text{and} \;
			\forall l. t \neq \bot^{l} 
		}
	}
	\end{array}
	\]

	Composition of middles \fbox{$ m \fatsemi^\ell (m,t) = (m,t) $}
	\[ 
	\begin{array}{rclclr}
	\comprule{\POOunit}{(\POOunit,t)}{
		(\POOunit,t)
	}{}
	\comprule{\POOfun{c_1}{c_2}}{(\POOfun{c_3}{c_4},t)}{
		(\POOfun{c_3 \fatsemi^{\ell} c_1}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{\POOprod{c_1}{c_2}}{(\POOprod{c_3}{c_4},t)}{
		(\POOprod{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{\POOsum{c_1}{c_2}}{(\POOsum{c_3}{c_4},t)}{
		(\POOsum{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4},t)
	}{}
	\comprule{m_1}{(m_2,t)}{
		(m_1,\bot^l)
	}{
		\sidecond{
			\ell = l \; \text{and} \;
			\neg m_1 \smile m_2 
		}
	}
	\end{array}
	\]

@@

	\fbox{$ cast(T,l,T) = c$}
	\[
	\begin{array}{rclr}
	\funrule{cast(T_1,l,T_2)}{
		id(T_1) \fatsemi^l id(T_2)
	}{}
	\end{array}
	\]
	\caption{\lazyD{} hypercoercion}
	\label{fig:HC-D}
\end{figure}

\figref{fig:HC-D} defines the constructors of \lazyD{} hypercoercion.

Let $ \ell $ range over $ \epsilon $ and labels.
$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

>>>>

\newcommand{\comprulel}[4]{#1 & \fatsemi^l & #2 & = & #3 & #4 \\}

@@

	\comprulel{
		\hyperCoercionI
	}{
		\hyperCoercionC{\epsilon}{m}{t}
	}{
		\hyperCoercionC{?^{l}}{m}{t}
	}{}

@@

	\comprule{
		\hyperCoercionC{h_1}{m_1}{t_1}
	}{
		\hyperCoercionC{\epsilon}{m_2}{t_2}
	}{
		\hyperCoercionC{h_1}{m_1 \fatsemi^{\ell} m_2}{t_2}
	}{
		\sidecond{
			m_1 \smile m_2
			\; \text{and} \;
			\forall l. t \neq \bot^{l}
		}
	}
	\comprule{
		\hyperCoercionC{h_1}{m_1}{t_1}
	}{
		\hyperCoercionC{?^l}{m_2}{t_2}
	}{
		\hyperCoercionC{h_1}{m_1 \fatsemi^{l} m_2}{t_2}
	}{
	\sidecond{
		m_1 \smile m_2
		\; \text{and} \;
		\forall l. t \neq \bot^{l}
	}
	}
	\comprulel{
	\hyperCoercionC{h_1}{m_1}{t_1}
	}{
	\hyperCoercionC{h_2}{m_2}{t_2}
	}{
	\hyperCoercionC{h_1}{m_1}{\bot^l}
	}{
	\sidecond{
		\neg \; m_1 \smile m_2
		\; \text{and} \;
		\forall l. t \neq \bot^{l}
	}
	}
%
%	\comprule{
%		\hyperCoercionC{h}{m_1}{t_1}
%	}{
%		\hyperCoercionC{\epsilon}{m_2}{t_2}
%	}{
%		\hyperCoercionC{h}{m'}{t'}
%	}{
%		\sidecond{
%			m_1 \fatsemi^{\ell} (m_2, t_2) = (m', t')
%			 \; \text{and} \;
%			 \forall l. t \neq \bot^{l}
%		}
%	}
%	
%	\comprule{
%		\hyperCoercionC{h}{m_1}{t_1}
%	}{
%		\hyperCoercionC{?^{l'}}{m_2}{t_2}
%	}{
%		\hyperCoercionC{h}{m'}{t'}
%	}{
%		\sidecond{
%			m_1 \fatsemi^{l'} (m_2, t_2) = (m', t')
%			\; \text{and} \;
%			\forall l. t \neq \bot^{l} 
%		}
%	}
	\end{array}
	\]

	Composition of middles \fbox{$ m \fatsemi^\ell m = m $}
	\[ 
	\begin{array}{rclclr}
	\comprule{\POOunit}{\POOunit}{
		\POOunit
	}{}
	\comprule{\POOfun{c_1}{c_2}}{\POOfun{c_3}{c_4}}{
		\POOfun{c_3 \fatsemi^{\ell} c_1}{c_2 \fatsemi^\ell c_4}
	}{}
	\comprule{\POOprod{c_1}{c_2}}{\POOprod{c_3}{c_4}}{
		\POOprod{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4}
	}{}
	\comprule{\POOsum{c_1}{c_2}}{\POOsum{c_3}{c_4}}{
		\POOsum{c_1 \fatsemi^\ell c_3}{c_2 \fatsemi^\ell c_4}
	}{}
	\end{array}
	\]

@@

	\fbox{$ cast(T,l,T) = c$}
	\[
	\begin{array}{rclr}
	\funrule{cast(T_1,l,T_2)}{
		id(T_1) \fatsemi^l id(T_2)
	}{}
	\end{array}
	\]
	\caption{\lazyD{} Hypercoercions}
	\label{fig:HC-D}
\end{figure}

\figref{fig:HC-D} defines functions that construct \lazyD{} hypercoercions. We 
expect language implementations to call $ id(T) $, which constructs an 
identity, $ seq(c,c) $, which composes two hypercoercions, and $ 
cast(T_1,l,T_2) $, which translate a cast to a hypercoercion. Other functions 
in the figure are their helpers.

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label. The last two lines are the most interesting. When the head of the 
second hypercoercion is a no-op, we use the same $ \ell $ to compose the 
middles. Otherwise, the head must be a projection, where we use the label in 
the projection to compose middles.

Let $ \ell $ range over $ \epsilon $ and labels.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.
